---
title: "HW1"
author: "Nihaal Rahman (nkr2121)"
due date: "2023-09-23"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(moderndive)
```

## Problem 1

Let's begin by loading in the data set. 

```{r}
data("early_january_weather")
```

This data set has `r nrow(early_january_weather)` observations of `r ncol(early_january_weather)` variables. The variables included in this data set are `r colnames(early_january_weather)`. The mean temperature recorded in early January 2013 was `r round(mean(early_january_weather$temp),2)` degrees.

Let's create a scatterplot of time vs. temperature. 
```{r}
ggplot(early_january_weather, aes(x=time_hour, y=temp, color=humid)) + geom_point()
```

We can see that the temperature steadily increased as the month progressed. Humidity was also low in the beginning of the month, and slowly increased as well. 

Now, I will save the scatter plot as a PNG file. 
```{r, message = FALSE}
ggsave('weather_scatter.png')
```

## Problem 2

Creating the dataframe: 

```{r}
set.seed(1234)

hw1_df = 
  tibble(
    sample = rnorm(n = 10),
    vec_logical = sample > 0,
    vec_character = c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A"),
    vec_factor = factor(vec_character)
  )
```

Now attempting to take the `mean` of `sample`, `vec_logical`, `vec_character`, and `vec_factor`: 
```{r}
mean(pull(hw1_df, sample))

mean(pull(hw1_df, vec_logical))

mean(pull(hw1_df, vec_character))

mean(pull(hw1_df, vec_factor))
```

As evident by the error messages we get for the last 2, you can take the mean of numeric or logical variables, but none of the other kinds of variables. 

Now, I will try applying the `as.numeric()` function to the `vec_logical`, `vec_character`, and `vec_factor` variables: 

```{r, results = FALSE, warning = FALSE}
as.numeric(pull(hw1_df, vec_logical))

as.numeric(pull(hw1_df, vec_character))

as.numeric(pull(hw1_df, vec_factor))
```

And then, will subsequently try taking the `mean` of these `as.numeric` conversions:

```{r, results = FALSE, warning = FALSE}
mean(as.numeric(pull(hw1_df, vec_logical)))

mean(as.numeric(pull(hw1_df, vec_character)))

mean(as.numeric(pull(hw1_df, vec_factor)))
```
The logical vector is able to be converted to a numeric variable by assigning 0's and 1's based on TRUE or FALSE status. The character vector is not able to be converted to a numeric variable, because the characters do not correspond to any numbers that R is able to read. However, the factor variable is able to be converted to a numeric variable, because the factors have corresponding levels, and those levels (3 factor levels in this case) can those levels are what R uses to convert the variable into a numeric variable. 